---
output: 
  bookdown::github_document2:
    number_figures: true
    number_sections: false
bibliography: inst/REFERENCES.bib
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dpeatdecomposition: R Interface to the Peatland Decomposition Database

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

`dpeatdecomposition` is an R interface to the Peatland Decomposition Database. The [Peatland Decomposition Database](---todo:link) [@Teickner.2024c] stores data from published litterbag experiments related to peatlands.

`dpeatdecomposition` uses functions from the [`dm`](https://github.com/cynkra/dm) and [`RMariaDB`](https://github.com/r-dbi/RMariaDB) package.

## Installation

You can install the development version of dpeatdecomposition like so:

``` r
remotes::install_github("henningte/dpeatdecomposition")
```

You will also need to download the [Peatland Decomposition Database](---todo) and open it in a running MariaDB instance.


## Example


```{r example, eval=FALSE}
library(dpeatdecomposition)

# connect to database
con <-
  RMariaDB::dbConnect(
    drv = RMariaDB::MariaDB(),
    dbname = "dpeatdecomposition",
    default.file = "~/my.cnf"
  )

# get database as dm object
dm_dpeatdecomposition <-
  dpeatdecomposition::dp_get_dm(con, learn_keys = TRUE)

# get information on samples
d <- 
  dm_dpeatdecomposition |>
  dm::pull_tbl(samples) |>
  as.data.frame()

# disconnect
RMariaDB::dbDisconnect(con)
```


# Contributions

If you want to contribute to the `dpeatdecomposition` package or the Peatland Decomposition Database, feel free top open an issue or contact me.


# Sources

```{r data-sources, echo=FALSE}
# connect to database
con <-
  RMariaDB::dbConnect(
    drv = RMariaDB::MariaDB(),
    dbname = "dpeatdecomposition",
    default.file = "~/my.cnf"
  )

# get database as dm object
dm_dpeatdecomposition <-
  dpeatdecomposition::dp_get_dm(con, learn_keys = TRUE)

# get information on samples
d <- 
  dm_dpeatdecomposition |>
  dm::pull_tbl(citations_to_datasets) |>
  as.data.frame()

# disconnect
RMariaDB::dbDisconnect(con)
```

Datasets currently included in the Peatland Decomposition Database are: `r d |> dplyr::mutate(id_citation = paste0("@", id_citation)) |> dplyr::pull(id_citation) |> paste(collapse = ", ")`


# Citation

Cite this package as: Henning Teickner, Klaus-Holger Knorr, and Edzer Pebesma (2024): dpeatdecomposition: R Interface to the Peatland Decomposition Database. ---todo: URL. accessed: `r Sys.Date()`. 

If you use data from the Peatland Decomposition Database, cite the database and the [sources](Sources) linked to each entry you use. For details, see [@Teickner.2024c]


# Acknowledgements

Development of this database was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) grant no. KN 929/23-1 to Klaus-Holger Knorr and grant no. PE 1632/18-1 to Edzer Pebesma.


# References
